---
title: "How the Immigration in Canada Affected the Voter Turnout Rate"
author: "Chien-Che Hung"
date: "10/14/2020"
output: pdf_document
---

```{r, include=FALSE}
library(readr)
library(brms)
library(tidyverse)
library(reshape2)
library(bayesplot)
library(ggplot2)
library(knitr)
gss <- read_csv("inputs/gss.csv")
```

```{r, include=FALSE}
gss$family_size <- as.character(gss$family_size)
gss$family_size[gss$family_size == "6"] <- "6 or more"

voting <- gss %>% filter(fed_elec_voted == "Yes" | fed_elec_voted == "No")
voting <- voting %>% filter(born_in_canada == "Born in Canada" | born_in_canada == "Born outside Canada")
```

```{r, eval = FALSE, include=FALSE}
# Set the weakly informed priors for the multilevel regression
prior_10 <- c(prior("normal(0,10)", class = "b", coef = "sense_belongNoopinion"),
           prior("normal(0,10)", class = "b", coef = "sense_belongNotstated"),
           prior("normal(0,10)", class = "b", coef = "sense_belongRefusal"),
           prior("normal(0,10)", class = "b", coef = "sense_belongSomewhatstrong"),
           prior("normal(0,10)", class = "b", coef = "sense_belongSomewhatweak"),
           prior("normal(0,10)", class = "b", coef = "sense_belongVerystrong"),
           prior("normal(0,10)", class = "b", coef = "sense_belongVeryweak"),
           prior("normal(0,10)", class = "b", coef = "born_in_canadaBornoutsideCanada"),
           prior("student_t(3, 0, 2.5)", class = sd, coef = "Intercept", group = "province"),
           prior("student_t(3, 0, 2.5)", class = sd, group = "province"))

# Fit the model by using the prior
sense_weak_fit <- brm(formula = fed_elec_voted ~ 1 + sense_belong + born_in_canada + (1|province), family = bernoulli(), seed = 123, data = voting, prior = prior_10, core = 4)
```


```{r, include=FALSE}
# load the already saved model
load("inputs/sense_weak_prior.rda")
```

#### Note:

- Here is the link to get the source code for cleaning the raw GSS data:
  - https://github.com/frankkhung/gss_data/blob/main/data_cleaning/gss_cleaning-1.R
  - Follow the instruction in the file to get the GSS 2013 Survey raw data
  - Download the files in this folder to get the dictionaries needed to tidy the data:
    - https://github.com/frankkhung/gss_data/tree/main/data_cleaning


- Here is the link to get the source code of this report:
  - https://github.com/frankkhung/gss_data/blob/main/data_analysis.Rmd
  - Please also download the files in this folder to make this file reproducible. 
    - https://github.com/frankkhung/gss_data/tree/main/inputs



# Abstract

As international immigration accounted for most of Canada's population growth, we want
to explore how this phenomenon affected civil participation in 2013. This report uses a
multilevel logistic regression model to analyze how a sense of belonging and birthplace
contribute to the voter turnout rate in the Federal election. As people born in Canada
would have different levels of sense of belonging than people born outside of Canada,
our analysis shows that a sense of belonging contributes a little to the voter turnout
rate. On the other hand, the single fact of whether the respondent born in Canada or not
negatively affected the voter turnout rate. Additionally, different provinces would
affect the voting participation rate.

# Introduction

This report is going to analyze the data obtained from the 2013 General Social Survey on
Social Identity. The survey topics range from social networks, civil participation,
birthplace to a sense of belonging and trust. Since there are varying topics that could
be covered, we will discuss whether the birthplace and sense of belonging would affect
the respondents voting turnout rate. Statistics Canada suggests that two-third of the
population growth was accounted for by international immigrants in 2006. The report from
CIC News points out that 82% of the population growth in 2019 came from immigration in
Canada (Thevenot, 2020). We want to explore whether the upsurge in immigration would affect the
participation rate in the Federal Election. The report structure includes explaining the
data, interpreting the model that we use, presenting the results, and, most importantly,
discussing the results.

# Data

This data covers varying topics, for example, social networks, civic participation, pride, the main activity of
respondent, birthplace, and well-being. With these categories, analysts can easily combine different variables to
conduct their desirable analysis. A meaningful analysis can be carried out with combinations of these results and
other background information and activities. The weakness of this data would be that there are 790 variables in total
from different categories. It would be tedious for analysts to go over each variable one by one or find the most
significant features from the dataset. Also, the data only includes respondents from ten provinces, excluding
residents of the Yukon, Northwest Territories, and Nunavut. By doing this, the survey and data would not be inclusive
enough. For this report, it considers variables that would be related to the election. The topic would be focusing on
the social participation in voting, including federal election, provincial election, and municipal election.

Several questions allow write-in responses, and it will be coded into a category in the question that fits the
response. By doing this, it is possible that the survey would lose some information or even categorize the response
into the wrong category. However, the advantage of the survey is relatively apparent. It touches on very detailed
information for the respondents. Also, since this survey is conducted through computer-assisted telephone
interviewing and electronic questionnaires, it is customizable for respondents who are not eligible for some
activities, such as voting.

This General Social Survey program surveys the ten provinces from June 2013 to March 2014, who were above
15-year-old. These surveys were asked through interviews via computer-assisted telephone interviewing (CATI) and
electronic (Internet) questionnaire (EQ). Unlike previous years' surveys conducted through Random Digit Dialing,
respondents in this 2013 survey were reached by telephone numbers available to Statistics Canada and the Address
Register. For Electronic Questionnaire, during the phone interview, if the respondents agreed to complete the survey
one, the interviewer would send them emails containing the survey and its access code.

As we mentioned above, the population was the ten provinces, which Yukon, Northwest Territories, and Nunavut were
excluded. For the frame for this survey, it was created from different components. The first component was the list
of telephone numbers to Statistics Canada from telephone companies or Census data. The second one was the Address
Register, which contains all dwellings in the ten provinces. For the sample size, the targeted sample size was 31973.
However, the number of respondents who answered was 27695.

# Model

Since we desire to explore whether the sense of belonging or respondents' birthplace would affectpeople's voting
turnout rate, our model's response variable would be binary. As in the Data Section mentioned, the respondents were
surveyed from different provinces (Stratum), which means that our observations might not be independent. As we have
repeat measures and heterogeneity across the provinces, we would be ignoring essential variations. To explore the
provinces' variance, we will use the Bayesian Multilevel (Hierarchical) Regression model to analyze the data. In the
multilevel model, the variances among observations are separated into Group-Level Effect and Population-Level
Effects. Group-Level Effect would be the variance across the provinces that GSS surveyed, and Population-Level
Effects would be the effects of the model's variables: Sense of Belonging and Born In Canada or Not. In other words,
we can say that this model is the compromise between the extremes:

1. No pooling: provinces pose no effects and independent on the voting turnout rate
2. Completely pooling: provinces have the same turnout rate

The Bayesian Multilevel Model utilizes the application of Bayes Theorem:

$$
\begin{aligned}
P(A|B) = \frac{P(B|A)P(A)}{P(B)}
\end{aligned}
$$

We can have the posterior estimation on the intercepts, coefficients, and variations because of the prior
information and the likelihood that we have on the response variable.

$$
\begin{aligned}
&P(\beta_{sense}, \beta_{born}, \alpha, \alpha_{prov},\sigma^2_{prov}|y) \\ &\propto P(y|\beta_{sense}, \beta_{born}, \alpha)P(\alpha_{prov}|\sigma^2_{prov})P(\beta_{sense})P(\beta_{born})P(\sigma^2_{prov})
\end{aligned}
$$

As we previously mentioned, in the model described below, we use a sense of belonging and whether born in Canada 
as fix intercepts and hierarchically model the effect of provinces.


$$
\begin{aligned}
Y_i|p_i & \sim Bernoulli(p_i), i = 1, ..., n\\
logit(p_i) = log(\frac{p_i}{1-p_i}) &= \alpha + \alpha_{province[i]} + \beta_{sense}belong + \beta_{born} born \\
\alpha_{province} & \sim student(0, 2.5) \\
\beta_{sense} & \sim Normal(0, 10) \\
\beta_{born} & \sim Normal(0, 10)
\end{aligned}
$$

For the priors that we pick are the weakly informed priors (Stan-Dev, 2020). As the coefficients in logistic regression could be a broad range, so we make the priors for the coefficients generally distributed with the variance of 10 and 0 mean. The standard deviation for the provinces would be in student-T distribution with 0 mean and 2.5 standard deviations. By having weakly informed prior, we can prevent our data from being too sensitive to our prior, especially with those provinces with fewer samples. 

To make sure this is a reasonable model, we want to check some diagnosis. Firstly, we want to see if the model converges. From the chart below (Graph 1),we can see that no divergences are showing in each estimate in different chains that we run in the model.

```{r, echo=FALSE, message=FALSE}
# plot the mcmc trace plot to check for the divergence of the model
mcmc_plot(sense_weak_fit, type = "trace") +
  labs(title = "Graph 1: Trace plot for Divergence") +
  theme(plot.title = element_text(face="bold"))
```


Secondly, we want to check if this plot has a proper fit. From the chart (Graph 2) we can see that the model has a remarkably similar density plot as the observations. Typically, for other Generalized Linear Models, one of the assumptions involves constant error variance. However, since the model we are using here is logistic regression and error variance is not a parameter in the Bernoulli distribution, we will not consider this assumption.

```{r, echo=FALSE, message=FALSE}
# plot the fitness of the model through pp_check
bayesplot::pp_check(sense_weak_fit) +
  labs(title ="Graph 2: Model Fitness")
```



The brms package runs this entire model and plots in R. It fits Bayesian generalized linear and
nonlinear multivariate multilevel models using 'Stan" for full Bayesian Inference
(https://cran.r-project.org/web/packages/brms/index.html). The following section will present and
discuss the results of this model.

# Result

Here are the general model and the fitness of the results.

```{r, echo=FALSE}
# show the summary of the model
summary(sense_weak_fit)
```


It presents the Group-Level (Province) Effects and Population-Level Effects (Sense of Belonging and
Born In Canada or Not). The estimation in the Population-Level Effects is shown in Log-Odds. If it is
positive, it means that it positively affects the odds and vice versa. With the assistance of the
distribution of each coefficient, we can see the importance of them. The ones that touch on zero 
indicate that there are some chances that they could be zero. Here, we can see that the intercept 
(respondents who did not state their sense of belonging level) and whether born in Canada from the 
Population-Level Effects are far from 0. It means that they have a certain degree of influence on the 
response variable.

Since we are assessing the effects in multilevel, we want to see how much variation that the province contributes to the model by finding the Intraclass Correlation Coefficient with the formula.
$$
\begin{aligned}
ICC = \frac{\sigma^2_{group}}{\sigma^2_{group}+\sigma^2_{population}}
\end{aligned}
$$

The variance for logistic distribution is $\frac{\pi^2}{3}$. From the result, we can say that provinces contribute 2.4% of the variance to the model.

```{r, echo=FALSE}
# calculate the icc 
tau2 <- VarCorr(sense_weak_fit)[[1]]$sd[1]^2
icc <- tau2 / (tau2 + (pi^2 / 3) )
icc <- as.data.frame(icc)
colnames(icc) <- "Probability"
rownames(icc) <- "Intraclass Correlation Coefficient"
# make icc into a table
kable(icc, digits = 3, caption = "ICC for province")
```

```{r, echo=FALSE}
# plot the posterior distribution for each posterior intercept estimate
mcmc_plot(sense_weak_fit, type = "areas") +
  labs(title = "Graph 3: Distribution Plot for Parameter Posterior Estimation")
```

From the broader view from the distribution graph, we can see that the provinces' variation will not be zero since the
distribution does not include zero, which means that provinces have effects on the election turnout. As we look closer
into the Group-Level Effects, the plot below (Graph 4) demonstrates the variations each province contributed.


```{r, echo=FALSE}
# get the group-level effect
random_effect <- ranef(sense_weak_fit)

# change the coefficients that we got above into a dataframe
r_Intercept <- random_effect$province[, , 1] %>%
  as_tibble() %>%
  rownames_to_column(var = "Provinces") %>%
  mutate(Variable = "sd(Intercept)")

# change the names in the r_Intercept data
province_names <- rownames(random_effect$province)
r_Intercept$Provinces <- province_names

# plot the variation graph for each 
r_Intercept %>%
  mutate(Contain_Zero = if_else(Q2.5*Q97.5 > 0, "no", "yes")) %>%
  ggplot(aes(x = Provinces, y = Estimate, col = Contain_Zero)) +
  geom_point() +
  geom_errorbar(aes(ymin=Q2.5, ymax=Q97.5)) +
  facet_grid(. ~ Variable, scale = "free") +
  coord_flip() +
  labs(title = "Graph 4: Variation in Each Province")
```

If the estimated variation credible interval contains zero, it means that the variation regarding the response variable
could be undetectable. On the other hand, if the credible interval does not contain zero, being in the province would
make a difference in the outcome.

We use logistic regression as our model to fit the binary response data. Due to this reason, we can calculate the mean
probability of each case by predicting each combination of cases and calculate the inverse logit scale. The below table
(Table 2) shows the probability of different cases. 

```{r, echo=FALSE}
# get the coefficient for the estimated coefficient
coeff <- as.data.frame(fixef(sense_weak_fit))

# calculate the log odds for each case (born in canada)
sense_dont_in <- coeff$Estimate[1] + coeff$Estimate[1]
sense_noop_in <- coeff$Estimate[1] + coeff$Estimate[2]
sense_notstate_in <- coeff$Estimate[1] + coeff$Estimate[3] 
sense_refuse_in <- coeff$Estimate[1] + coeff$Estimate[4] 
sense_somestrong_in <- coeff$Estimate[1] + coeff$Estimate[5] 
sense_someweak_in <- coeff$Estimate[1] + coeff$Estimate[5]
sense_strong_in <- coeff$Estimate[1] + coeff$Estimate[6] 
sense_weak_in <- coeff$Estimate[1] + coeff$Estimate[7]
odd_in <- as.data.frame(rbind(sense_dont_in, sense_noop_in, sense_notstate_in, sense_refuse_in, sense_somestrong_in, sense_someweak_in, sense_strong_in, sense_weak_in))

# calculate the log odds for each case (born outside of canada)
sense_dont_out <- coeff$Estimate[1] + coeff$Estimate[9]
sense_noop_out <- coeff$Estimate[1] + coeff$Estimate[2] + coeff$Estimate[9]
sense_notstate_out <- coeff$Estimate[1] + coeff$Estimate[3] + coeff$Estimate[9]
sense_refuse_out <- coeff$Estimate[1] + coeff$Estimate[4] + coeff$Estimate[9]
sense_somestrong_out <- coeff$Estimate[1] + coeff$Estimate[5] + coeff$Estimate[9]
sense_someweak_out <- coeff$Estimate[1] + coeff$Estimate[5] + coeff$Estimate[9]
sense_strong_out <- coeff$Estimate[1] + coeff$Estimate[6] + coeff$Estimate[9]
sense_weak_out <- coeff$Estimate[1] + coeff$Estimate[7] + coeff$Estimate[9]
odd_out <- as.data.frame(rbind(sense_dont_out, sense_noop_out, sense_notstate_out, sense_refuse_out, sense_somestrong_out, sense_someweak_out, sense_strong_out, sense_weak_out))

# combine the tables above
odds <- rbind(odd_in, odd_out)

# change the odds data by inverse logit
odds$V1 <- 1/(1+exp(-odds$V1))

# change the column name
colnames(odds) <- "Probability"

# present the table
kable(odds, caption = "Mean Probability for each case", digits = 3)
```


# Discussion

## Result Discussion

Graph 3, even though some of the credible intervals for coefficient estimate include 0,
the coefficients have the possibility of posing no effects to the voting turnout.
However, we will still take a look at the effects of each of them. In the table below,
we can see that the respondents who did not know about their feeling regarding the sense
of belonging to Canada positively influence the log-odds. Here we define odds as in the
ratio between the amounts staked by the parties to a bet. People who did not have any
opinion, somewhat strong opinion, and somewhat weak opinion have a negative relationship
with the log-odds of the probability of voting in the Federal Election. 

```{r, echo=FALSE}
# change intercept into "dont know" for sense of belonging
rownames(coeff)[1] <- "sense_belongDontknow"
# create the table for the coefficients
kable(coeff, digits = 3, caption = "Coefficient Estimates")
```

Noticeably, the respondents born outside of Canada pose the most substantial adverse
effect on the Federal Election voting turnout. It means that people born outside of
Canada, residing in Canada while the survey was conducted, did not vote during the 
election. The below table shows the probability of voting in each situation with the 
mean probability of them. We can see that people born in Canada have a way higher 
probability of voting inthe Federal Election. (TABLE 4)

```{r, echo=FALSE}
# present the probability for each case (separated by born in or out canada)
prob_sense <- as.data.frame(cbind(odds$Probability[1:8], odds$Probability[9:16]))
rownames(prob_sense) <- c("sense_dont", "sense_noop", "sense_notstate", "sense_refuse", "sense_somestrong", "sense_someweak", "sense_strong", "sense_weak")
colnames(prob_sense) <- c("Born In", "Born Out")
mean <- as.data.frame(cbind(mean(odds$Probability[1:8]), mean(odds$Probability[9:16])))
colnames(mean) <- c("Born In", "Born Out")
prob_sense <- rbind(prob_sense, mean)
rownames(prob_sense)[9] <- "Mean Probability"
kable(prob_sense, caption = "Mean Probability for each case", digits = 3)
```


Interestingly, as we look at people born outside of Canada's response to the sense of
belonging, most of them responded in the "Strong" category. (Table 5) This result
indicates that there are underlying reasons why this group of people has negative
effects on voting.

```{r, echo=FALSE}
# explore the count for sense belong's categories for respondents born outside of canada
voting <- voting %>% filter(born_in_canada == "Born outside Canada")
out_belong <- as.data.frame(count(voting, sense_belong))
kable(out_belong, caption = "Counts for Sense of Belonging Response (Born Outside of Canada)")
```

Graph 4 shows the variation caused in the Group-Level (provinces). Specifically, residents in Quebec, Saskatchewan, and Alberta would be most likely to vote while living in Ontario, Newfoundland, Labrador, New Brunswick, Manitoba, and British Columbia could suggest being in these provinces would not be the incentives of why the residences vote.

To conclude the findings in a broader context, a sense of belonging to Canada possibly would not contribute to voting turnout. On the other hand, the respondents' birthplace creates a massive difference between whether people would vote. For different provinces in Canada, some of the provinces would influence the decision of whether their residents should vote or not.

## Weakness and Caveats

There are some possible weaknesses and caveats in the model. From the analysis above, we
can see that most of the answers from the sense of belonging question could not affect
the voting turnout. It could also be the weakly informed prior has too many effects on
the sense of belonging variable. Even though we could get a considerable amount of
information from this data and analysis, this research topic is limited to Canada and
the Federal Election. Specifically, the data is collected by a stratified simple random
sample, and each stratum has a different situation, as we have shown above. It would be
hard and infeasible to generalize this conclusion regarding a sense of belonging to
other countries. However, as immigration is happening in other countries, the massive
difference in whether born-in-Canada could be utilized in other analyses.

## Future Work

POLYAS, an online voting company, states that there is an increasing division between
different voting groups. One evidence is that in General Election in the UK in 2015, 75%
of the upper-middle class and the middle class voted, whereas only 56% of the working
class and the non-working class voted. (Increase Voter Turnout: Create a Sense of Belonging 2017)
This evidence from POLYAS shows that the
salary level or education level in this GSS 2013 Social Identity Survey could contribute
to the probability of voting turnout rate. If we only conclude from the analysis that
this report has done, a sense of belonging would not be a significant social
participation factor. On the other hand, there is a significant disparity between
whether the respondents were born in Canada. We could dig more into this group of
people, and policymakers could also shift their ideologies towards the people born
outside of Canada to increase the voter turnout rate.

# Reference

## Data Citation

### Data
- Statistics Canada (2013). Canadian General Social Surveys Cycle 27: Social Identity 2013 (Version 2)

### Data Cleaning Code Adapted From:
- Alexander R. and Caetano S. (2020) gss_clearning-1.R (Version 1) 

## Report Citation

- Gabry J, Mahr T (2020). “bayesplot: Plotting for Bayesian
  Models.” R package version 1.7.2, <URL:
  https://mc-stan.org/bayesplot>.
- Hadley Wickham, Jim Hester and Romain Francois (2018). readr:
  Read Rectangular Text Data. R package version 1.3.1.
  https://CRAN.R-project.org/package=readr
- Hadley Wickham (2007). Reshaping Data with the reshape
  Package. Journal of Statistical Software, 21(12), 1-20. URL
  http://www.jstatsoft.org/v21/i12/.
- H. Wickham. ggplot2: Elegant Graphics for Data Analysis.
  Springer-Verlag New York, 2016.
- Increase Voter Turnout: Create a Sense of Belonging. (2017, January 27). Retrieved October 19, 2020, from   https://www.polyas.com/increase-voter-turnout/sense-of-belonging
- Immigration and growth. Statistics Canada: Canada's national statistical agency / Statistique Canada : Organisme statistique national du Canada. (2016, October 7). https://www150.statcan.gc.ca/n1/pub/11-402-x/2010000/chap/imm/imm01-eng.htm. 
- JJ Allaire and Yihui Xie and Jonathan McPherson and Javier
  Luraschi and Kevin Ushey and Aron Atkins and Hadley Wickham
  and Joe Cheng and Winston Chang and Richard Iannone (2020).
  rmarkdown: Dynamic Documents for R. R package version 2.3. URL
  https://rmarkdown.rstudio.com.
- Paul-Christian Bürkner (2017). brms: An R Package for Bayesian
  Multilevel Models Using Stan. Journal of Statistical Software,
  80(1), 1-28. doi:10.18637/jss.v080.i01
- Paul-Christian Bürkner (2018). Advanced Bayesian Multilevel
  Modeling with the R Package brms. The R Journal, 10(1),
  395-411. doi:10.32614/RJ-2018-017
- R Core Team (2020). R: A language and environment for
  statistical computing. R Foundation for Statistical Computing,
  Vienna, Austria. URL https://www.R-project.org/.
- Stan-Dev. stan-dev/stan. GitHub. https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations. 
- Thevenot, S. (2020, July 20). Immigration to drive Canada's population growth as global birth rates fall. CIC News. https://www.cicnews.com/2020/07/immigration-to-drive-canadas-population-growth-as-global-birthrates-fall-0715086.html. 
- Wickham et al., (2019). Welcome to the tidyverse. Journal of
  Open Source Software, 4(43), 1686,
  https://doi.org/10.21105/joss.01686
- Yihui Xie (2020). knitr: A General-Purpose Package for Dynamic
  Report Generation in R. R package version 1.29.
